### 2025-05-02 docs/Homeownership_BankOfMomAndDad.tex
Followed Will feedback

I then study how illiquidity affects the strategic behavior in the absence of commitment. I find that 27\% of young households prefer illiquid housing---due to sales costs---to liquid housing. These households have wealthy parents and use the sales costs as a commitment device to maintain low liquid wealth in the next period, effectively becoming ``wealthy hand-to-mouth.'' Lower liquid wealth implies higher marginal propensities to consume (MPC) and, hence, more parental transfers. As an illustration, consider a child who buys a house and brings no liquid wealth to the next period. If the parent does not transfer in the next period, the child will either face very low consumption---implying  a high marginal value of wealth---or be forced to liquidate the house and incur sales costs. The parent dislikes both outcomes and may find it optimal to transfer enough to keep the child in the home.

That parental wealth generates a preference for illiquidity is not only a new theoretical mechanism---distinct from explanations like temptation or time inconsistency \citep[e.g.,][]{attanasio2024temptation,laibson1997golden}---but one that helps explain the positive link between parental wealth and successfull homeownership documented earlier. It also aligns with \citet{choukhmane2023benefits}, who find that households with wealthy parents are more likely to invest in illiquid retirement accounts and avoid costly early withdrawals, even after controlling for a wide set of characteristics.

I then examine the implications of transfers and housing on marginal propensities to consume (MPCs). First, the presence of altruistic transfers reduces average MPCs by nearly 30\%, suggesting that standard life-cycle models without parent-child interactions may overstate MPCs. Second, the preference for illiquidity induced by transfers offers a novel explanation for the existence of high-MPC households \cite[e.g.,][]{kaplan2022marginal}; in fact, the entire decline in average MPCs is driven by lower MPCs among the wealthy hand-to-mouth. Third, a common explanation for high MPCs is heterogeneity in preferences---such as impatience \citep[e.g.,][]{aguiar2024hand}---but such behavior is observationally similar to that of households with wealthy, altruistic parents.

>>>>

I then study how illiquidity affects strategic behavior in the absence of commitment. As an illustration, consider a child who buys a house and brings no liquid wealth to the next period. If the parent does not transfer in the next period, the child will either face very low consumption—implying a high marginal value of wealth—or be forced to liquidate the house and incur sales costs. The parent dislikes both outcomes and may find it optimal to transfer enough to keep the child in the home. More precisely, children with wealthy parents use the sales costs as a commitment device to maintain low liquid wealth in the next period, raising their next-period marginal utility of wealth and thereby strengthening the parent’s transfer motive. As a result, 27\% of young households prefer illiquid housing over liquid housing. That parental wealth generates a preference for illiquidity is a new theoretical mechanism---distinct from existing explanations such as temptation or time inconsistency \citep[e.g.,][]{attanasio2024temptation,laibson1997golden}.

This lower cost of illiquidity for households with wealthy parents offers a natural explanation for the positive empirical relationship between parental wealth and successful homeownership I found in the PSID. It also aligns with \citet{choukhmane2023benefits}, who find that households with wealthy parents are more likely to invest in illiquid retirement accounts and avoid costly early withdrawals, even after controlling for a wide set of characteristics.

I then examine the interplay  of transfers and housing for marginal propensities to consume (MPC). A substantial literature has sought to reconcile structural models with the high MPCs observed in empirical data.  In my model, the introduction of altruistic transfers lowers average MPCs by about 30\%, implying that standard life-cycle models may understate the discrepancy.  Moreover, one common explanation for high MPCs is that some households are wealthy but liquidity constrained---such as the ``wealthy hand-to-mouth'' in \cite{kaplan2022marginal}---but I find that ignoring parental transfers overstates their MPCs by about 50\%. Another explanation for high MPCs is heterogeneity in preferences---such as impatience \citep[e.g.,][]{aguiar2024hand}---yet such behavior is observationally similar to that of households with wealthy, altruistic parents.
